Carbonate chemistry at Ny-Ålesund

### 2021-04-06 awipev-CO2.Rmd
Minor changes

## Impact K1K2  {.tabset .tabset-pills}
### Lueker vs Sulpis
```{r impact of constants discrete, echo=FALSE, warning=FALSE, message=FALSE}
# Compare Lueker and Sulpis using mean discrete values
# first, order columns
d_all <- d_all %>% 
  dplyr::select(order(colnames(d_all))) %>% 
  dplyr::select(datetime, date, year, monthd, montht, everything())
# calculate means of discrete values
mean_ct <- mean(discrete$ct, na.rm = TRUE)
mean_at <- mean(discrete$at, na.rm = TRUE)
mean_pco2 <- mean(discrete$pco2_calc_fb, na.rm = TRUE)
mean_sal <- mean(discrete$sal_fb, na.rm = TRUE)
mean_temp <- mean(discrete$temp_fb, na.rm = TRUE)
mean_discrete <- tibble(mean_ct = mean_ct, mean_at = mean_at, mean_pco2 = mean_pco2, mean_sal = mean_sal, mean_temp = mean_temp)

# calculate typical uncertainty assuming uncertainties of 3 uatm for pCO2 and 4 umol/kg for AT and CT
errors_pco2_at <- seacarb::errors(flag=24, var1=mean_pco2, var2=1e-6*mean_at, S=mean_sal,
                          T=mean_temp, P=0, Pt=0, Sit=0, k1k2="l", kf="dg", ks="d", pHscale="T",
                          b="u74", evar1=3, evar2 = 4e-6)
errors_at_ct <- seacarb::errors(flag=15, var1=1e-6*mean_at, var2=1e-6*mean_ct, S=mean_sal,
                          T=mean_temp, P=0, Pt=0, Sit=0, k1k2="l", kf="dg", ks="d", pHscale="T",
                          b="u74", evar1=4e-6, evar2 = 4e-6)
errors <- tibble(e_pH = c(errors_pco2_at$pH, errors_at_ct$pH),
                 e_pCO2 = c(NA, errors_at_ct$pCO2),
                 e_DIC = c(errors_pco2_at$pH, NA)*1e3,
                 e_Oa = c(errors_pco2_at$OmegaAragonite, errors_at_ct$OmegaAragonite)
)
# impacts using pco2 and AT pair
carb_l_1 <- carb(24, mean_pco2, mean_at*1e-6, S = mean_sal, T = mean_temp, P = 0, 
                 Pt = 0, Sit = 0,
                 k1k2 = "l", kf = "dg", ks = "d", pHscale = "T", b = "u74")
carb_s_1 <- carb(24, mean_pco2, mean_at*1e-6, S = mean_sal, T = mean_temp, P = 0, 
                 Pt = 0, Sit = 0,
                 k1k2 = "s20", kf = "dg", ks = "d", pHscale = "T", b = "u74")
diff_ls_1 <- carb_l_1 - carb_s_1
# impacts using AT-CT pair
carb_l_2 <- carb(15, mean_at*1e-6, mean_ct*1e-6, S = mean_sal, T = mean_temp, P = 0, 
                 Pt = 0, Sit = 0,
                 k1k2 = "l", kf = "dg", ks = "d", pHscale = "T", b = "u74")
carb_s_2 <- carb(15, mean_at*1e-6, mean_ct*1e-6, S = mean_sal, T = mean_temp, P = 0,
                Pt = 0, Sit = 0, 
                k1k2 = "s20", kf = "dg", ks = "d", pHscale = "T", b = "u74")
diff_ls_2 <- carb_l_2 - carb_s_2
# summary
const_choice <- bind_rows(diff_ls_1, diff_ls_2) %>% 
  dplyr::mutate(flag = c("pCO2 and ALK", "ALK and DIC"),
                S = mean_sal,
                T = mean_temp,
                DIC = DIC * 1e6
                ) %>% 
  dplyr::select(flag, pH, pCO2, DIC, OmegaAragonite) %>% 
  dplyr::bind_cols(errors)
```
Using the following using mean discrete values:
`r knitr::kable(mean_discrete, digits = c(1, 1, 0, 1, 1), caption = "Mean discrete values.", format = "simple")`

@@

# save elements for k1k2 report
k1k2_pco2_at <- list(mean_discrete = mean_discrete, const_choice = const_choice, tmp2 = tmp2,
             errors_pco2_at = errors_pco2_at, diff_table_pco2_at = diff_table_pco2_at, 
             diff_ls_quant = diff_ls_quant)
save(k1k2_pco2_at, file="reports/k1k2_pco2_at.RData")
```

The discrete values of pCO2, AT, salinity and temperature in the FerryBox were used to calculate pH using the formulations for K1 and K2 of Lueker et al. (2000) and Sulpis et al. (2020). The differences induced by the choice of the carbonic acid constants are:

@@

# save elements for k1k2 report
k1k2_at_ct <- list(errors_at_ct = errors_at_ct,
                   diff_table_at_ct = diff_table_at_ct)
save(k1k2_at_ct, file="reports/k1k2_at_ct.RData")
```
Here we compare pH measured spectrophotometrically with pH calculated from AT and CT using various formulations of K1 and K2. All pH values were normalized to a temperature of 4°C. The absolute differences are generally large, up to `r round(max(abs(diff_table_at_ct[, 2:5])), digits=3)` pH units. In general, all formulations overestimate spectrophotometric pH. pH calculated using the formulation of Papadimitriou et al. (2018) is closer to measured pH, with a mean difference of `r round(diff_table_at_ct[4, 4], digits=3)` pH units. This difference is `r abs(round(diff_table_at_ct[4, 4]/errors_at_ct$pH, digits=1))` times the uncertainty for pH calculated from AT and CT according to seacarb (`r round(errors_at_ct$pH, digits=3)`). The differences found with the other formulations are, on average, slightly lower than the uncertainty for pH calculated from AT and CT according to seacarb.

>>>>

## K1K2  {.tabset .tabset-pills}
```{r impact of constants discrete, echo=FALSE, warning=FALSE, message=FALSE}
# Compare Lueker and Sulpis using mean discrete values
# first, order columns
d_all <- d_all %>% 
  dplyr::select(order(colnames(d_all))) %>% 
  dplyr::select(datetime, date, year, monthd, montht, everything())
# calculate means of discrete values
mean_ct <- mean(discrete$ct, na.rm = TRUE)
mean_at <- mean(discrete$at, na.rm = TRUE)
mean_pco2 <- mean(discrete$pco2_calc_fb, na.rm = TRUE)
mean_sal <- mean(discrete$sal_fb, na.rm = TRUE)
mean_temp <- mean(discrete$temp_fb, na.rm = TRUE)
mean_discrete <- tibble(mean_ct = mean_ct, mean_at = mean_at, mean_pco2 = mean_pco2, mean_sal = mean_sal, mean_temp = mean_temp)

# calculate typical uncertainty assuming uncertainties of 3 uatm for pCO2 and 4 umol/kg for AT and CT
errors_pco2_at <- seacarb::errors(flag=24, var1=mean_pco2, var2=1e-6*mean_at, S=mean_sal,
                          T=mean_temp, P=0, Pt=0, Sit=0, k1k2="l", kf="dg", ks="d", pHscale="T",
                          b="u74", evar1=3, evar2 = 4e-6)
errors_at_ct <- seacarb::errors(flag=15, var1=1e-6*mean_at, var2=1e-6*mean_ct, S=mean_sal,
                          T=mean_temp, P=0, Pt=0, Sit=0, k1k2="l", kf="dg", ks="d", pHscale="T",
                          b="u74", evar1=4e-6, evar2 = 4e-6)
errors <- tibble(e_pH = c(errors_pco2_at$pH, errors_at_ct$pH),
                 e_pCO2 = c(NA, errors_at_ct$pCO2),
                 e_DIC = c(errors_pco2_at$pH, NA)*1e3,
                 e_Oa = c(errors_pco2_at$OmegaAragonite, errors_at_ct$OmegaAragonite)
)
# impacts using pco2 and AT pair
carb_l_1 <- carb(24, mean_pco2, mean_at*1e-6, S = mean_sal, T = mean_temp, P = 0, 
                 Pt = 0, Sit = 0,
                 k1k2 = "l", kf = "dg", ks = "d", pHscale = "T", b = "u74")
carb_s_1 <- carb(24, mean_pco2, mean_at*1e-6, S = mean_sal, T = mean_temp, P = 0, 
                 Pt = 0, Sit = 0,
                 k1k2 = "s20", kf = "dg", ks = "d", pHscale = "T", b = "u74")
diff_ls_1 <- carb_l_1 - carb_s_1
# impacts using AT-CT pair
carb_l_2 <- carb(15, mean_at*1e-6, mean_ct*1e-6, S = mean_sal, T = mean_temp, P = 0, 
                 Pt = 0, Sit = 0,
                 k1k2 = "l", kf = "dg", ks = "d", pHscale = "T", b = "u74")
carb_s_2 <- carb(15, mean_at*1e-6, mean_ct*1e-6, S = mean_sal, T = mean_temp, P = 0,
                Pt = 0, Sit = 0, 
                k1k2 = "s20", kf = "dg", ks = "d", pHscale = "T", b = "u74")
diff_ls_2 <- carb_l_2 - carb_s_2
# summary
const_choice <- bind_rows(diff_ls_1, diff_ls_2) %>% 
  dplyr::mutate(flag = c("pCO2 and ALK", "ALK and DIC"),
                S = mean_sal,
                T = mean_temp,
                DIC = DIC * 1e6
                ) %>% 
  dplyr::select(flag, pH, pCO2, DIC, OmegaAragonite) %>% 
  dplyr::bind_cols(errors)
```

@@

# save elements for k1k2 report
k1k2_pco2_at <- list(mean_discrete = mean_discrete, const_choice = const_choice, tmp2 = tmp2,
             errors_pco2_at = errors_pco2_at, diff_table_pco2_at = diff_table_pco2_at, 
             diff_ls_quant = diff_ls_quant)
save(k1k2_pco2_at, file="reports/k1k2_pco2_at.RData")
```
#### Difference Lueker vs Sulpis
Using the following using mean discrete values:
`r knitr::kable(mean_discrete, digits = c(1, 1, 0, 1, 1), caption = "Mean discrete values.", format = "simple")`

#### Difference between all formulations and measured pH
The discrete values of pCO2, AT, salinity and temperature in the FerryBox were used to calculate pH using the formulations for K1 and K2 of Lueker et al. (2000) and Sulpis et al. (2020). The differences induced by the choice of the carbonic acid constants are:

@@

# save elements for k1k2 report
k1k2_at_ct <- list(errors_at_ct = errors_at_ct,
                   diff_table_at_ct = diff_table_at_ct)
save(k1k2_at_ct, file="reports/k1k2_at_ct.RData")
```
#### Difference Lueker vs Sulpis
Using the following using mean discrete values:
`r knitr::kable(mean_discrete, digits = c(1, 1, 0, 1, 1), caption = "Mean discrete values.", format = "simple")`

#### Difference between all formulations and measured pH
Here we compare pH measured spectrophotometrically with pH calculated from AT and CT using various formulations of K1 and K2. All pH values were normalized to a temperature of 4°C. The absolute differences are generally large, up to `r round(max(abs(diff_table_at_ct[, 2:5])), digits=3)` pH units. In general, all formulations overestimate spectrophotometric pH. pH calculated using the formulation of Papadimitriou et al. (2018) is closer to measured pH, with a mean difference of `r round(diff_table_at_ct[4, 4], digits=3)` pH units. This difference is `r abs(round(diff_table_at_ct[4, 4]/errors_at_ct$pH, digits=1))` times the uncertainty for pH calculated from AT and CT according to seacarb (`r round(errors_at_ct$pH, digits=3)`). The differences found with the other formulations are, on average, slightly lower than the uncertainty for pH calculated from AT and CT according to seacarb.
